### 2019-02-27 2-Chapters/4-Chapter/chapter4.tex
Done for now.

In Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the many ``intelligent'' end-devices.

@@

We presented in Section~\ref{sec:4:gnuradio} a demonstration, showed in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
%
By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.

@@

In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.

@@

The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
We prefer to consider that all devices have the same probability to transmit to keep the notation as simple as possible.

Moreover, for sake of simplicity we supposed that all devices use the same standard. Our future work will consider more realistic interference scenarios and IoT networks, with, \emph{e.g.}, non-slotted time, more than one base station etc.

Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \emph{e.g.}, one transmission a day.
So additionally to deciding the channel for communication, each device has to also decide when to communicate.
% This is another direction of research, that we will investigate in the future.

@@

Possible future extensions of our demonstration include the following points.
We could consider more dynamic objects (\emph{e.g.}, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing new hardware.
%
% Pub pour notre autre article
% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.

@@

The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works. Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
%
For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
that nevertheless have not been explored for the targeted problem.
Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this section.

>>>>

In Section~\ref{sec:4:firstModel}, we proposed an evaluation of the performance of MAB learning algorithms in IoT networks,
with a focus on the convergence of algorithms, in terms of successful transmission rates, when the proportion of intelligent dynamic devices changes.
Concretely, increasing this probability allows to insert more objects in the same network, while maintaining a good Quality of Service.
We show that \UCB{} and TS have near-optimal performance, even when their underlying \iid{} assumption is violated by the presences many ``intelligent'' end-devices which follow a random activation process.

@@

We presented in Section~\ref{sec:4:gnuradio} a demonstration, showed in $2018$ at the ICT conference \cite{Besson2018ICT}, and further detailed in the companion paper \cite{Besson2019WCNC}.
We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
%
By using such low-cost algorithms, we demonstrated empirically that a dynamically re-configurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels, by using the acknowledgement (\Ack) feedback sent from the base station.

@@

In Section~\ref{sec:4:retransmissions}, we presented an extension of our model of LPWA networks based on an ALOHA protocol, slotted both in time and frequency, in which dynamic IoT devices can again use machine learning algorithms, to improve their Packet Loss Ratio (PLR) when accessing the network.
The main novelty of this model is to address the packet retransmissions upon radio collision, by using a Multi-Armed Bandit framework.
We presented and evaluated several learning heuristic that try to learn how to transmit and retransmit in a smarter way, by using the \UCB{} algorithm for channel selection for first transmission, and different proposals based on \UCB{} for the retransmissions upon collisions.

@@

The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.

Moreover, for sake of simplicity we supposed that all devices use the same standard.
Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.

Another extension could be to not have a Bernoulli process (or any random process), but a fixed rate of transmission, \eg, one transmission a day.
So additionally to deciding the channel for communication (\ie, \emph{where} to communicate), each device has to also decide \emph{when} to communicate.
% This is another direction of research, that we will investigate in the future.
However, this clearly leads to a much larger action space, as there are many time slots in one day (for example), and thus we believe that as soon as the action space becomes too large in this extension, the simple MAB-based learning approach could be no longer appropriate.
It is well-known in the MAB literature that the larger the action space, the slower is the convergence speed of any stationary MAB algorithms.
It could be interesting to study the possible application of \emph{contextual} MAB \cite{Li10,Luo17} or structured MAB \cite{Combes17} models and algorithms for this extension.

@@

Possible future extensions of our demonstration include the following points.
We could consider more dynamic objects (\eg, $100$) but it would either cost more in terms of equipment, or in terms of software engineering to simulation more objects with the same card.
We could also implement a real-world IoT communication protocol (like the LoRaWAN standard), which we prefer not to do as it would cost a significant effort of development.
Finally, we could also study the interference in case of other gateways located nearby, and this could be done without needing a lot of new hardware (using one extra USRP card is enough to simulate another gateway).
%
% Pub pour notre autre article
% We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.

@@

The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works. Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
%
For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
that nevertheless have not been explored for the targeted problem.
Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this Chapter.
